GEE models

### 2017-10-15 doc/esm.Rmd
expanded on pls multivariable vs gee
added more detail about DAG

For the covariates, they were chosen based on previous literature, from 
directed acyclic graph (DAG) recommendations, and from QIC. The DAG
recommendations were obtained from using the DAGitty software,
http://dagitty.net/. DAG is a technique that allows for identifying potential
confounding causal pathways and to limit potential bias introduced from
adjusting for colliding variables. A collider variable is a variable that is
determined from two or more other variables, such that in causal graph modelling,
the arrows from the other variables appear to 'collide' into the variable. QIC 
is a technique that balances the goodness of fit of the model with its relative
complexity (i.e. number of covariates) against other models being compared. This
identifies a specific model from a number of models that is the 'best' fit and
least complex of all the models. QIC and other information criterion methods are
used in conjunction with other techniques such as literature and DAG based. The 
final GEE model selected as best fitting was negligibly different when using 
insulin sensitivity or beta-cell function as an outcome and as such the same 
covariates were chosen for models with each of the outcomes. The covariates
medication use, WC, ALT, and MET would be considered classical confounders while
TAG would be considered a mediating variable.

Partial least squares (PLS) regression is a technique similar to principal
component analysis (PCA) that is designed to extract meaningful information from
multivariate, high dimensionality data (e.g. as in metabolomic and other -omic type 
analyses). These types of methods try to extract as much of the variance in the 
data into a smaller number of components or factors. The difference between PCA 
and PLS is that PLS is a supervised method, while PCA is not. This means that
PLS uses a response variable(s), i.e. an outcome or **Y**, to 'supervise' (or
constrain) the variation inherent in the predictors (**X**) while PCA only
describes the variation inherent in **X**. Because of this, PLS can be better at
predicting the contributions of predictor variables against an outcome variable.

>>>>

For the covariates, they were chosen based on previous literature, from directed
acyclic graph (DAG) recommendations, and from QIC. The DAG recommendations were
obtained from using the DAGitty software, http://dagitty.net/. DAG is a
technique that allows for identifying potential confounding causal pathways and
to limit potential bias. DAGs are composed on nodes (variables) connected by
edges (arrows; causal pathway between two variables). Using DAGs to identify
variables to adjust for done algorithmically, as implemented in the DAGitty
software. DAGitty calculates possible sets of variables to include in a model to
adjust for potential confounding in the causal network. Use of DAGs in this
manner can identify potential sources of bias, such as from inclusion of
colliding variables. A collider variable is a variable that has two (or more)
'parent' nodes/variables (connected causally by DAG edges pointing into the
'child' node, which is the collider). The collider variable and parent variables
are directed linked to the exposure and the outcome. Including all variables
involved with the collider could introduce bias as it does not generate an
unconditional association between all involved variables in the model.

QIC is a technique that balances the goodness of fit of the model with its
relative complexity (i.e. number of covariates) against other models being
compared. This identifies a specific model from a number of models that is the
'best' fit and least complex of all the models. QIC and other information
criterion methods are used in conjunction with other techniques such as
literature and DAG based. The final GEE model selected as best fitting was
negligibly different when using insulin sensitivity or beta-cell function as an
outcome and as such the same covariates were chosen for models with each of the
outcomes. The covariates medication use, WC, ALT, and MET would be considered
classical confounders while TAG would be considered a mediating variable.

Partial least squares (PLS) regression is a technique similar to principal
component analysis (PCA) that is designed to extract meaningful information from
multivariate, high dimensionality data (e.g. as in metabolomic and other -omic type 
analyses). These types of methods try to extract as much of the variance in the 
data into a smaller number of components or factors. The difference between PCA 
and PLS is that PLS is a supervised method, while PCA is not. This means that
PLS uses a response variable(s), i.e. an outcome or **Y**, to 'supervise' (or
constrain) the variation inherent in the predictors (**X**) while PCA only
describes the variation inherent in **X**. Because of this, PLS can be better at
predicting the contributions of predictor variables against an outcome variable.
In this analysis, PLS was computed on the NEFA as a proportion due to the
inherent interdependence of the values (all must equal 100%). This 
interdependence allows PLS to identify patterns in the composition, since
the concentration data is all highly positively correlated with each other.
